Model Checking Differentially Private Properties

### 2018-05-02 examples.tex
add description of subsampling

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. The range of
the mechanism
however is $\bbfZ$. It may give nonsensical outputs such as
negative integers for non-negative queries.

The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
  \ldots, m \}$}
outputs $f (x) + Z$ where $Z$ is a random variable with the following
distribution:
\[
\Pr[Z = z] =
\left\{
  \begin{array}{ll}
    0 & \textmd{ if } z < - f (x) \\
    \frac{\alpha^{f (x)}}{1 + \alpha} & \textmd{ if } z = -f (x)\\
    \frac{1 - \alpha}{1 + \alpha}\alpha^{|z|} &
    \textmd{ if } -f (x) < z < m - f (x)\\
    \frac{\alpha^{m - f (x)}}{1 + \alpha} & \textmd{ if } z = m-f (x)\\
    0 & \textmd{ if } z > m - f (x)
  \end{array}
\right.
\]
Note the range of the truncated $\alpha$-geometric mechanism is
$\{ 0, 1, \ldots, m \}$. The truncated $\alpha$-geometric mechanism is
again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$.
The truncated $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
given in Figure~\ref{figure:geometric-mechanism-table}.

@@

\subsection{Subsampling}
\label{subsection:subsampling}

>>>>

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
%Observe that $\Pr[Y=0] \rightarrow 1$ as $\alpha \rightarrow 0$. The
%$\alpha$-geometric mechanism is less noisy for smaller $\alpha$.
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. Observe that
the privay guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
sensitivity of the query $f$. To achieve $(\epsilon, 0)$-differential
privacy using the $\alpha$-geometric mechanism, one first decides the
sensitivity of the query and then computes the parameter $\alpha =
e^{-{\epsilon}/{{\Delta (f)}}}$.

The range of the mechanism is $\bbfZ$. It may give nonsensical
outputs such as negative integers for non-negative queries.
The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
  \ldots, m \}$}
outputs $f (x) + Z$ where $Z$ is a random variable with the 
distribution:
\[
\Pr[Z = z] =
\left\{
  \begin{array}{ll}
    0 & \textmd{ if } z < - f (x) \\
    \frac{\alpha^{f (x)}}{1 + \alpha} & \textmd{ if } z = -f (x)\\
    \frac{1 - \alpha}{1 + \alpha}\alpha^{|z|} &
    \textmd{ if } -f (x) < z < m - f (x)\\
    \frac{\alpha^{m - f (x)}}{1 + \alpha} & \textmd{ if } z = m-f (x)\\
    0 & \textmd{ if } z > m - f (x)
  \end{array}
\right.
\]
Note the range of the truncated $\alpha$-geometric mechanism is
$\{ 0, 1, \ldots, m \}$. The truncated $\alpha$-geometric mechanism is
again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$.
The truncated $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
given in Figure~\ref{figure:geometric-mechanism-table}.

@@

\subsection{Subsampling Majority}
\label{subsection:subsampling}

### 2018-05-02 subsampling.tex
add description of subsampling

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{Subsampling}{$d$, $f$}
    \Require{$d \in \{ R, B \}^3$, $f : \{ R, B \}^3 \rightarrow \{ R,
      B \}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \State{$z \leftarrow \{ f (\hat{d}_1), f (\hat{d}_2), \ldots, f
      (\hat{d}_m) \}$}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
    \Else
      \State{\Return {$\bot$}}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling}
  \label{algorithm:subsampling}
\end{algorithm}

@@

      \node[node] (top) at (2.5,  0.75) { $\mathit{out}$ };
      \node[node] (bot) at (2.5, -0.75) { $\bot$ };

@@

      \end{tikzpicture}
    }
  \caption{Markov Chain for Subsampling}
  \label{figure:mc-subsampling}
\end{figure}

>>>>

The sensitivity of queries is required to apply the (truncated)
$\alpha$-geometric mechanism. Recall that the sensitivity of a query
is the maximal difference of query results on any two neighbors. There are
two problems for such mechanisms depending on query sensitivity. First, 
sensitivity of queries can be hard to compute or estimate. Second,
the sensitivity over arbitrary neighbors can be too conservative for
the actual dataset in use. One therefore would like to have
mechanisms independent of query sensitivity.

Subsampling is a technique to design such mechanisms. Concretely, let
us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
is the majority in the dataset while respecting individual
privacy. This can be achieved as follows. The mechanism first samples $m$
sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
$d$. It then computes $m$ query results from the $m$
sub-datasets. Finally, the mechanism reports a perturbed result in a
privacy respecting way
(Algorithm~\ref{algorithm:subsampling-majority}).

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SubsamplingMajority}{$d$, $f$}
    \Require{$d \in \{ R, B \}^n$, $f : \{ R, B \}^* \rightarrow \{ R,
      B \}$}
    \State{$q, m \leftarrow 
            \frac{\epsilon}{64\ln (1/\delta)},
            \frac{\log (n/\delta)}{q^2}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
    \Else
      \State{\Return {$\bot$}}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling Majority}
  \label{algorithm:subsampling-majority}
\end{algorithm}

@@

      \node[node] (top) at (2.5,  0.75) { $\top$ };
      \node[node] (bot) at (2.5, -0.75) { $\bot$ };

@@

      \end{tikzpicture}
    }
  \caption{Markov Chain for Subsampling Majority}
  \label{figure:mc-subsampling}
\end{figure}
